Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2021-12-17 src/ms.tex
jvs comments

The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}). Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff. As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages. Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale. Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge. These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.

@@

Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models. In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e. the low-mass stars would be rotating more slowly than the model predictions). Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law; the models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$. A higher normalization factor would cause the low-mass stars to spin down more at fixed age.

Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models. At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence. The core and envelope at these times are thus assumed to be decoupled. However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015}. When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age. This hypothesis is discussed further in \S\ref{subsec:shortperiod}.

>>>>

The location of the short-period pile-up is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}). Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff. As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages. Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale. Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge. These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.

@@

Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models. In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e. the low-mass stars would be rotating more slowly than the model predictions). Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law. The models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$. A higher normalization factor would cause the low-mass stars to spin down more at fixed age.

Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models. At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence. The core and envelope at these times are thus assumed to be decoupled. However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}. When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age. This hypothesis is discussed further in \S\ref{subsec:shortperiod}.
